---
title: "The relationship between international study and civic virtues"
author:
  - name: Yangyue Li
    corresponding: true
    email: yangyueli28@uchicago.edu
    affiliations:
      - id: id1
        name: "University of Chicago"
#        group: Developmental Psychology Program
        department: MA Program in the Social Sciences
        address: "1155 E 60th St."
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
abstract: This study investigated civic virtues among international and American undergraduate students to examine potential differences in psychological constructs associated with civic development. we compared 86 students from a single university across multiple measures of civic virtues, including relational civility, civic attitudes and behaviors, empathy, epistemic humility, and cultural competence. Independent samples t-tests and one-way analyses of variance found no statistically significant differences between international and American students across these constructs. The findings suggest that the contemporary university experience may promote comparable civic development trajectories, regardless of students' international status. 
keywords: [civic virtues, empathy, wisdom, civility, international studies]
# If true, tables and figures are mingled with the text instead of listed at the end of the document.
floatsintext: true
# Numbered lines (.pdf and .docx only)
numbered-lines: false
# File with references
bibliography: bibliography.bib
# Suppress title page
suppress-title-page: false
# Link citations to references
link-citations: true
# If true, adds today's date below author affiliations. If text, can be any value.
# This is not standard APA format, but it is convenient.
# Works with docx, html, and typst. 
draft-date: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  title-impact-statement: "Impact Statement"
  references-meta-analysis: "References marked with an asterisk indicate studies included in the meta-analysis."
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    prefer-html: true
    documentmode: man
    keep-tex: true
    # You can turn these off for a simple document
    # But you probably want them for a thesis or chapter
    list-of-figures: true
    list-of-tables: true
    # Change default output size of figures unless specified in the chunk
    fig-width: 2
    fig-height: 2
    out-width: 40%
---

```{r}
#| label: setup
#| include: false

# install.packages("ggcorrplot")
# install.packages("ggpubr")

library(stats)
library(psych)
library(ggpubr)
library(ggcorrplot)
library(conflicted)
library(tidyverse)
library(flextable)
library(ftExtra)
library(knitr)
library(kableExtra) 
library(rempsyc) 
conflicts_prefer(dplyr::filter, .quiet = TRUE)
conflicts_prefer(flextable::separate_header, .quiet = TRUE)

# Read the cleaned data
# I cleaned my raw data in scripts/clean_data.R
civic_data <- read_csv("civic-virtues-dataset/civic.clean.csv")
```

```{r}
#| label: analysis-preferences
#| include: false

# Set seed for reproducibility
set.seed(42)

# Set global chunk options
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed,
  fig.height = 6,
  fig.width = 7,
  message = FALSE,
  warning = FALSE,
  comment = ""
)
```

In an interconnected world, civic virtues are crucial for fostering responsible citizenship and encouraging individuals to prioritize societal well-being over self-interest [@sherrodDimensionsCitizenshipOpportunities2002]. U.S. universities, as diverse microcosms, bring domestic and international students together, offering unique contexts for civic engagement and understanding the values of others. Domestic American students may develop civic engagement through strong ties to their local communities and the American education system, while international students may bring perspectives shaped by cross-cultural experiences and adaptation to a new cultural environment. Undergoing the same admission process and sharing similar college life on campus with their American peers, international students may still differ in civic behaviors. These contexts raise critical questions about the factors that foster civic virtues and motivate individuals to engage in civic-minded behaviors. How do experiences in a foreign and culturally different environment influence civic virtues and related psychological characteristics such as empathy and cultural competence? Do international students differ on these psychological aspects from US students?

## Literature Review

### Education and Civic Virtues

Civic engagement may foster a sense of belonging, purpose, and responsibility within a community, and may empower individuals to contribute to a common good [@flanaganCivicEngagementTransition2010]. Research suggests that civic education - educational experiences that are directed at increasing civic virtues - happens mostly during adolescence, but can continue into young adulthood [@flanaganCivicEngagementTransition2010; @sherrodDimensionsCitizenshipOpportunities2002]. Much of this experience occurs in educational settings, particularly colleges, which offer structured opportunities through coursework, extracurricular activities, community service programs, and student organizations [@flanaganCivicEngagementTransition2010]. These experiences could help young adults develop civic virtues and social responsibility.

Besides formal instruction in character education and civility [@jeynesMetaAnalysisRelationshipCharacter2019; @torney-purtaSchoolsRoleDeveloping2002], non-didactic and informal experiences have important implications for the development of civic virtues as well. Indeed, research highlights that youths who are involved in community-based organizations or extracurricular activities are more likely to be civically active [@zarrettRoleOrganizedActivities2021]. Furthermore, a study suggests that participating in community service or volunteering is associated with adolescent civic beliefs about considering similar types of civic engagement behaviors [@metzgerAdolescentsCivicEngagement2019]. Similarly, @vezinaInvestigatingCivicParticipation2019 found that prosocial/community-based activities are related to a greater likelihood of being in the high-sustained civic engagement trajectory.

Additionally, recent research suggests that study abroad programs may increase some aspects of civic virtues and related psychological capacities. Living and studying in a foreign cultural environment provide opportunities for students to interact with people from diverse backgrounds, fostering skills related to civic virtues such as empathy, open-mindedness, and perspective-taking in a non-didactic way [@chieffoLargeScaleAssessmentStudent2004]. Research indicates that even short-term study abroad programs enhance student’s empathy and open-mindedness toward diverse perspectives, and foster a deeper sense of responsibility as global citizens [@chieffoLargeScaleAssessmentStudent2004]. @blackAssessingImpactBusiness2006 assessed the impact of a month-long business-focused study abroad program in London on students' cultural awareness and personal development. Consistent with previous findings, students showed significant improvements in cross-cultural empathy and understanding of global perspectives. A more recent study examined the relationship between studying abroad and civic virtues [@boulwareStrangerStrangeLand2023]. The results suggest that undergraduate students who have gone abroad via one particular set of study abroad programs in undergraduate studies demonstrated higher levels of empathy, civic engagement, and civility toward others, compared with students who do not have these experiences or have no interest in them.

### International Students and Civic Virtues

Universities in the United States enroll a substantial number of international students annually, yet the benefits for this population remain underexplored in respect of non-curricular benefits. Research on a variety of different psychological capacities wisdom, bilingualism, and cultural adaptation may offer insights into the interaction of personal and contextual factors that could shape civic virtues among international students. As @staudingerWhatPredictsWisdomrelated1998 point out, virtues are a fundamental component of wisdom, with openness to experience and social competence serving as critical predictors of wisdom-related performance [@baltesFascinationWisdomIts2008]. Similarly, [@grossmannWisdomContext2017] illustrates how contextual factors, such as cultural and experiential environments, shape wisdom-related reasoning and behaviors. International students, who frequently navigate diverse cultural and educational settings, may adapt their civic engagement to align with situational demands. 

## Current Study

Although international students represent a substantial portion of the US college population and can play an important role in civic life, there is little research on how this experience relates to civic virtues. This study aims to investigate civic virtues among international students studying in the United States. The primary goals are to answer the key research question: do international students, due to their unique experiences in a foreign country, exhibit similar or differing levels of civic virtues compared to American students? Given that study abroad experiences can positively influence one’s civility, we hypothesize that international students may exhibit higher levels of civic virtues than American students with no experience studying abroad. This would be comparable to US students who study abroad suggesting that living in a foreign country may be an important experience related to civic virtues. Alternatively, international students may score higher on some psychological characteristics related to civic virtue development, including epistemic humility, empathy, need for cognition (NFC), and cultural competence, while showing comparable levels of civic attitudes and behaviors.

# Method

## Participants

```{r}
#| label: demographics

# Create a new dataset for demographic analysis
demo_data <- civic_data %>%
  mutate(`Student Status` = ifelse(`Student Status` == "American.sa", "American", `Student Status`)) %>%
  select(`Student Status`, Age, Gender, Ethnicity)

# Calculate Age Mean and SD
age_summary <- demo_data %>%
  group_by(`Student Status`) %>%
  summarise(Age = sprintf("%.2f (%.2f)", mean(Age, na.rm = TRUE), sd(Age, na.rm = TRUE))) %>%
  mutate(Label = "Age (years)", Level = "Mean (SD)") %>%
  pivot_wider(names_from = `Student Status`, values_from = Age)

# Calculate Gender Counts and Percentages
gender_summary <- demo_data %>%
  group_by(`Student Status`, Gender) %>%
  summarise(n = n(), .groups = 'drop') %>%
  group_by(`Student Status`) %>%
  mutate(Percentage = round(n / sum(n) * 100, 1)) %>%
  mutate(Value = sprintf("%d (%.1f%%)", n, Percentage)) %>%
  select(`Student Status`, Gender, Value) %>%
  pivot_wider(names_from = `Student Status`, values_from = Value) %>%
  rename(Level = Gender) %>%
  mutate(Label = "Gender") %>%
  arrange(Level)  # Ensure Gender subcategories are ordered

# Ensure "Gender" appears only once in the first column
gender_summary$Label[duplicated(gender_summary$Label)] <- ""

# Calculate Ethnicity Counts and Percentages
ethnicity_summary <- demo_data %>%
  group_by(`Student Status`, Ethnicity) %>%
  summarise(n = n(), .groups = 'drop') %>%
  group_by(`Student Status`) %>%
  mutate(Percentage = round(n / sum(n) * 100, 1)) %>%
  mutate(Value = sprintf("%d (%.1f%%)", n, Percentage)) %>%
  select(`Student Status`, Ethnicity, Value) %>%
  pivot_wider(names_from = `Student Status`, values_from = Value) %>%
  rename(Level = Ethnicity) %>%
  mutate(Label = "Ethnicity") %>%
  arrange(Level)  # Ensure Ethnicity subcategories are ordered

# Ensure "Ethnicity" appears only once in the first column
ethnicity_summary$Label[duplicated(ethnicity_summary$Label)] <- ""

# Combine all sections into a single table
demographic_table <- bind_rows(age_summary, gender_summary, ethnicity_summary)

```

```{r}
#| label: tbl-demographics
#| tbl-cap: Demographic Characteristics of Participants

demographic_table %>%
  kable(align = "l", col.names = c("Label", "Level", "American", "International")) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

A survey was adapted from Boulware et al. (2023). Participants were recruited through the Psychology Department’s SONA system during the winter quarter in 2025 as well as through campus advertisements. The total sample consisted of `r nrow(civic_data)` participants, with `r sum(civic_data[["Student Status"]] == "International")` international students and `r sum(civic_data[["Student Status"]] %in% c("American", "American.sa"))` domestic students, representing a diverse composition of student backgrounds as detailed in @tbl-demographics.

## Measures

A survey was adapted from Boulware et al. (2023), and distributed through the Psychology Department’s research participant recruitment system during the winter quarter in 2025.

### Civic Virtues

We will use two measures to assess civic virtues. The first is Doolittle and Faul’s (2013) 14-item Civic Engagement Scale. It includes the 8 civic attitude subscale (e.g., “I feel responsible for my community”, 1 = disagree, and 7 = agree) and 6 civic behavior items (e.g., “I am involved in structured volunteer position(s) in the community”, 1 = never, and 7 = always). The second one is based on the Workplace Relational Civility scale, tailored for student population. It includes 13 items assessing civility towards others and 13 items assessing civility experienced by participants (Di Fabio & Gori, 2016). Sample items include “I was able to express my values and my beliefs calmly to others” and “Others were able to express their point of view without being disrespectful toward me” (1= not at all, and 5 = a great deal).

### Other constructs related to civic virtues

We will use a scale developed at the Center for Practical Wisdom at the University of Chicago to assess epistemic humility (Hoeckner, 2011, personal communication). Sample items include “I accept that many good and bad things happening to me are beyond my control” (1 = strongly disagree, and 8 = strongly agree). We will use the 8-item Empathy Quotient to measure empathy towards others [@loewenEightitemFormEmpathy]. Sample items include “I find it easy to put myself in somebody else’s shoes” (1 = never, and 5 = always). We will use the 6-item NFC short form scale to measure the tendency to engage in cognitive effort [@linsdeholandacoelhoVeryEfficientAssessment2020]. Sample questions include “I would prefer complex to simple problems” (1 = extremely uncharacteristic, and 5 = extremely characteristic). We will measure cultural competency based on a set of questions provided by the Study Abroad program at the University of Chicago. For international students, we will have questions asking them to report their first language, English proficiency (indicated by the Test of English as a Foreign Language or a self-report scale), and duration of their residence in the United States. American students will indicate whether they have studied abroad.

## Data Analysis

To test the first hypothesis, independent t-tests compared international and American students on each measure of civic virtues (civic attitudes, civic behaviors, and civility) and other related constructs. These comparisons indicated whether international students, as a group, exhibit different levels of civic virtues compared to their American peers. An Analysis of Variance (ANOVA) will be conducted to account for relationships among the dependent variables and provide a more comprehensive analysis.

# Results

## Descriptive Analyses

```{r}
#| label: prepare-descriptive-data
#| code-summary: "Prepare data for descriptive statistics"

# Select only numeric columns for analysis
desc_numeric_data <- civic_data %>%
  select(where(is.numeric), -`subject #`, -`CC.sa`, -`Age`)
```

```{r}
#| label: summary-function
#| code-summary: "Basic summary of variables"
#| code-fold: true
#| include: false

# Provide summary of numeric variables
summary(desc_numeric_data)
```

```{r}
#| label: tbl-descriptive-stats
#| tbl-cap: Descriptive Statistics of Civic Virtues Measures

# Calculate descriptive statistics
desc_stats <- desc_numeric_data %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarize(
    Mean = mean(Value, na.rm = TRUE),
    SD = sd(Value, na.rm = TRUE),
    Median = median(Value, na.rm = TRUE),
    Min = min(Value, na.rm = TRUE),
    Max = max(Value, na.rm = TRUE)
  ) %>%
  mutate(across(c(Mean, SD, Median, Min, Max), 
                ~round(., 2)))

kbl(desc_stats, 
    booktabs = TRUE) 
```

```{r}
#| label: summarize_construct
#| eval: false
#| include: false

# Define a function to summarize a construct across student groups
summarize_construct <- function(data, construct) {
  data %>%
    group_by(`Student Status`) %>%
    summarise(
      Mean = mean(.data[[construct]], na.rm = TRUE),
      SD = sd(.data[[construct]], na.rm = TRUE),
      N = n()
    ) %>%
    arrange(desc(Mean))  # Sort by highest mean
}

summarize_construct(civic_data, "Civic Attitudes")

summarize_construct(civic_data, "Civility Total")
```


Descriptive statistics for the key measures of civic virtues are presented in @tbl-descriptive-stats. The sample exhibited varying levels of psychological constructs related to civic engagement. *Civic Attitudes* showed a mean of `r desc_stats$Mean[desc_stats$Variable == "Civic Attitudes"]` (*SD* = `r desc_stats$SD[desc_stats$Variable == "Civic Attitudes"]`), with a range from `r desc_stats$Min[desc_stats$Variable == "Civic Attitudes"]` to `r desc_stats$Max[desc_stats$Variable == "Civic Attitudes"]`. *Civic Behaviors* displayed a mean of `r desc_stats$Mean[desc_stats$Variable == "Civic Behaviors"]` (*SD* = `r desc_stats$SD[desc_stats$Variable == "Civic Behaviors"]`), suggesting moderate levels of community engagement.

The variability across different civic virtue measures suggests potential differences in how participants conceptualize and enact civic engagement. *Epistemic Humility*, for instance, ranged from `r desc_stats$Min[desc_stats$Variable == "Epistemic Humility"]` to `r desc_stats$Max[desc_stats$Variable == "Epistemic Humility"]`, with a mean of `r desc_stats$Mean[desc_stats$Variable == "Epistemic Humility"]` (*SD* = `r desc_stats$SD[desc_stats$Variable == "Epistemic Humility"]`), indicating considerable individual variation in this construct.

## Bivariate Correlations

```{r}
#| label: bivariate-correlations

# Select constructs for correlation analysis
correlation_data <- civic_data %>%
  select(
    `Civility Total`,
    `CB_CA Total`, 
    `Wise Reasoning`,
    `Need for Cognition`,
    `Empathy`,
    `Need for Closure`,
    `Epistemic Humility`,
    `Cultural Competence`
  )

# Function to compute correlation matrix with p-values

corr.test <- function(x) {
  # Number of variables
  n <- ncol(x)
  
  # Create matrices to store correlation and p-values
  r <- matrix(NA, n, n)
  p <- matrix(NA, n, n)
  
  # Compute correlations and p-values
  for (i in 1:n) {
    for (j in 1:n) {
      if (i != j) {
        test <- cor.test(x[,i], x[,j])
        r[i,j] <- round(test$estimate, 3)
        p[i,j] <- round(test$p.value, 3)
      } else {
        r[i,j] <- 1
        p[i,j] <- NA
      }
    }
  }
  
  # Add column and row names
  colnames(r) <- colnames(x)
  rownames(r) <- colnames(x)
  colnames(p) <- colnames(x)
  rownames(p) <- colnames(x)
  
  return(list(r = r, p = p))
}

# Compute correlation matrix
corr_results <- corr.test(as.matrix(correlation_data))

# Create correlation matrix with significance markers
corr_matrix_formatted <- matrix("", nrow(corr_results$r), ncol(corr_results$r))
rownames(corr_matrix_formatted) <- rownames(corr_results$r)
colnames(corr_matrix_formatted) <- colnames(corr_results$r)

for (i in 1:nrow(corr_results$r)) {
  for (j in 1:ncol(corr_results$r)) {
    if (i != j) {
      # Add significance markers
      sig_marker <- if (corr_results$p[i,j] < 0.001) "***" 
                    else if (corr_results$p[i,j] < 0.01) "**" 
                    else if (corr_results$p[i,j] < 0.05) "*" 
                    else "ns"
      
      corr_matrix_formatted[i,j] <- paste0(
        round(corr_results$r[i,j], 3), 
        sig_marker
      )
    } else {
      corr_matrix_formatted[i,j] <- "1.000"
    }
  }
}

# Convert correlation matrix to a data frame
corr_table <- corr_matrix_formatted %>%
  as.data.frame() %>%
  rownames_to_column("Variable")
```

```{r}
#| label: tbl-correlations
#| tbl-cap: Bivariate Correlations among Civic Virtues Constructs
# Generate correlation table 

corr_table %>%
kbl(format = "latex",
    booktabs = TRUE) %>%
  kable_styling(
    latex_options = c("scale_down", "landscape", "hold_position")
  )
```

```{r}
#| label: fig-correlation-plot
#| fig-cap: "Correlation Matrix of Civic Virtues Constructs"
#| fig-height: 7
#| fig-width: 7

ggcorrplot::ggcorrplot(
  corr_results$r,  # Use numeric correlation matrix instead
  type = "lower",  
  show.diag = FALSE,  
  lab = TRUE,  
  lab_size = 3,  
  method = "circle",  
  colors = c("#6D9EC1", "white", "#E64B35"),  
  title = "Correlation Matrix of Civic Virtues Constructs"
)
```

The correlation matrix presented in @tbl-correlations and visualized in @fig-correlation-plot provides a comprehensive view of these intricate relationships. *CB_CA Total* was moderately correlated with *Wise Reasoning* (r = `r corr_results$r[2,3]`, p = `r corr_results$p[2,3]`), suggesting an interconnection between behavioral and cognitive aspects of civic engagement. *Empathy* showed notable correlations with several constructs. It was moderately correlated with *Wise Reasoning* (r = `r corr_results$r[5,3]`, p = `r corr_results$p[5,3]`), and significantly associated with *Epistemic Humility* (r = `r corr_results$r[5,7]`, p = `r corr_results$p[5,7]`). These relationships highlight the potential interconnectedness of psychological characteristics related to civic virtues. *Cultural Competence* demonstrated interesting associations, including a moderate correlation with *Need for Cognition* (r = `r corr_results$r[8,4]`, p = `r corr_results$p[8,4]`). The significant correlations, marked with asterisks in @tbl-correlations, indicate the complex interplay between different psychological constructs related to civic engagement.[^1]

[^1]: Significance levels are indicated as follows: *p* < 0.05, **p** < 0.01, ***p*** < 0.001.

## T-tests

```{r}
#| label: t-test-analysis

# Prepare data for t-tests
t_test_data <- civic_data %>%
  mutate(
    StudentGroup = if_else(`Student Status` %in% c("American", "American.sa"), 
                           "American", 
                           "International")
  ) %>%
  select(
    StudentGroup,
    `Civility Total`,
    `CB_CA Total`, 
    `Wise Reasoning`,
    `Need for Cognition`,
    `Empathy`,
    `Need for Closure`,
    `Epistemic Humility`,
    `Cultural Competence`
  )

# Perform t-tests for each construct
perform_t_test <- function(data, variable, group) {
  # Separate groups
  american_group <- data %>% 
    filter(StudentGroup == "American") %>% 
    pull(!!variable)
  international_group <- data %>% 
    filter(StudentGroup == "International") %>% 
    pull(!!variable)
  
  # Perform t-test with equal variances
  test_result <- t.test(american_group, international_group, var.equal = TRUE)
  
  # Return formatted results
  tibble(
    Variable = variable,
    t_statistic = round(test_result$statistic, 3),
    df = round(test_result$parameter, 0),  # Now an integer
    p_value = round(test_result$p.value, 3),
    mean_american = round(mean(american_group, na.rm = TRUE), 3),
    sd_american = round(sd(american_group, na.rm = TRUE), 3),
    mean_international = round(mean(international_group, na.rm = TRUE), 3),
    sd_international = round(sd(international_group, na.rm = TRUE), 3)
  )
}

# Run t-tests for all selected constructs
t_test_results <- map_dfr(
  c("Civility Total", "CB_CA Total", 
    "Wise Reasoning", "Need for Cognition", "Empathy", 
    "Need for Closure", "Epistemic Humility", "Cultural Competence"),
  ~perform_t_test(t_test_data, .x, "StudentGroup")
)
```

```{r}
#| label: tbl-t-tests
#| tbl-cap: Comparison of Civic Virtues Constructs between American and International Students

t_test_results %>%
  kbl(
    booktabs = TRUE,
    col.names = c("Variable", "t", "df", "p", 
                  "M (American)", "SD (American)", 
                  "M (International)", "SD (International)")
  ) %>%
  kable_styling(
    latex_options = c("scale_down", "landscape", "hold_position")
  )

```

```{r}
#| label: fig-t-test-viz
#| fig-cap: "Comparison of Civic Virtues Constructs between American and International Students"
#| fig-height: 12
#| fig-width: 15

plot_data <- civic_data %>%
  mutate(
    StudentGroup = if_else(`Student Status` %in% c("American", "American.sa"), 
                           "American", 
                           "International")
  ) %>%
  pivot_longer(
    cols = c(`Civility Total`, `CB_CA Total`, 
             `Wise Reasoning`, `Need for Cognition`, `Empathy`, 
             `Need for Closure`, `Epistemic Humility`, `Cultural Competence`),
    names_to = "Variable",
    values_to = "Value"
  )

# Create a custom labeller function
custom_labels <- c("CB_CA Total" = "Civic Behaviors and Civic Attitudes",
                   "Civility Total" = "Relational Civility")

ggplot(plot_data, aes(x = StudentGroup, y = Value, fill = StudentGroup)) +
  geom_boxplot(
    alpha = 0.7,  
    width = 0.7
  ) +
  
  stat_compare_means(
    method = "t.test",
    label = "p.signif",
    label.y.npc = 0.95
  ) +
  
  # Facet wrap to create separate panels
  facet_wrap(~ Variable, scales = "free_y", ncol = 3, labeller = labeller(Variable = custom_labels)) +
  
  scale_fill_manual(values = c("American" = "#FFB6C1", "International" = "#8FBC8F")) +
  
  
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),                 # Rotate x-axis labels
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),
    plot.subtitle = element_text(hjust = 0.5, size = 15),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold")
  ) +
  
  labs(
    title = "Comparison of Civic Virtues Constructs",
    subtitle = "American vs International Students",
    x = "Student Group",
    y = "Score",
    fill = "Student Group"
  ) 

```

```{r}
#| label: t-test-cacb
#| include: false

t_test_ca <- t.test(
  civic_data$`Civic Attitudes`[civic_data$`Student Status` %in% c("American", "American.sa")],
  civic_data$`Civic Attitudes`[civic_data$`Student Status` == "International"],
  var.equal = TRUE  # Use FALSE if variances are unequal
)

# Print the t-test results
print(t_test_ca)

t_test_cb <- t.test(
  civic_data$`Civic Behaviors`[civic_data$`Student Status` %in% c("American", "American.sa")],
  civic_data$`Civic Behaviors`[civic_data$`Student Status` == "International"],
  var.equal = TRUE  # Use FALSE if variances are unequal
)

# Print the t-test results
print(t_test_cb)
```

```{r}
#| label: fig-civic-behaviors-attitudes
#| fig-cap: "Differences in Civic Behaviors and Civic Attitudes Between International and American Students"
#| fig-width: 8
#| fig-height: 5

# Prepare data for visualization
civic_data_filtered <- civic_data %>%
  filter(`Student Status` %in% c("American", "American.sa", "International")) %>%
  mutate(`Student Status` = ifelse(`Student Status` %in% c("American", "American.sa"), "American", "International")) %>%
  select(`Student Status`, `Civic Behaviors`, `Civic Attitudes`) %>%
  pivot_longer(cols = c(`Civic Behaviors`, `Civic Attitudes`), names_to = "Measure", values_to = "Score")

# Create the enhanced boxplot with separate y-axis for each facet
ggplot(civic_data_filtered, aes(x = `Student Status`, y = Score, fill = `Student Status`)) +
  geom_boxplot(alpha = 0.5, color = "black", outlier.shape = NA, width = 0.6) +  
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.8, aes(color = `Student Status`)) +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 4, color = "black", stroke = 1.2) + 
  facet_wrap(~ Measure, scales = "free_y") +  # ✅ NOW WORKS! Separate y-axis per variable
  scale_fill_manual(values = c("American" = "#1b9e77", "International" = "#d95f02")) +
  scale_color_manual(values = c("American" = "#1b9e77", "International" = "#d95f02")) +
  labs(
    title = "Civic Behaviors & Attitudes: International vs. American Students",
    subtitle = "Boxplots with jittered data points and mean indicators (Separate y-axis)",
    x = "Student Status", y = "Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    strip.text = element_text(face = "bold", size = 14)
  )

```

```{r}
#| label: fig-civility-me-others
#| fig-cap: "Differences in Civility Towards Me and Civility Towards Others Between International and American Students"
#| fig-width: 8
#| fig-height: 5

civility_data_filtered <- civic_data %>%
  filter(`Student Status` %in% c("American", "American.sa", "International")) %>%
  mutate(`Student Status` = ifelse(`Student Status` %in% c("American", "American.sa"), "American", "International")) %>%
  select(`Student Status`, `Civility Towards Me`, `Civility Towards Others`) %>%
  pivot_longer(cols = c(`Civility Towards Me`, `Civility Towards Others`), names_to = "Measure", values_to = "Score")

# Create the enhanced boxplot with facet_grid
ggplot(civility_data_filtered, aes(x = `Student Status`, y = Score, fill = `Student Status`)) +
  geom_boxplot(alpha = 0.5, color = "black", outlier.shape = NA, width = 0.6) +  
  geom_jitter(width = 0.15, alpha = 0.5, size = 1.8, aes(color = `Student Status`)) +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 4, color = "black", stroke = 1.2) + 
  facet_grid(. ~ Measure) +
  scale_fill_manual(values = c("American" = "#7570b3", "International" = "#e7298a")) +
  scale_color_manual(values = c("American" = "#7570b3", "International" = "#e7298a")) +
  labs(
    title = "Civility Towards Me & Others: International vs. American Students",
    subtitle = "Boxplots with jittered data points and mean indicators",
    x = "Student Status", y = "Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    strip.text = element_text(face = "bold", size = 14)
  )
```

The dataset includes `r sprintf("a total of %d student groups", length(unique(civic_data[["Student Status"]])))`. However, in this study, we only compared American students with international students. To examine differences between American students (including those with study abroad experience) and international students, a series of independent samples t-tests were conducted across key civic and civility constructs. The results are summarized in @tbl-t-tests and @fig-t-test-viz. No siginificant differences were found in all of these constructs. 
An independent samples t-test was conducted to examine differences in Civic Attitudes (CA) and Civic Behaviors (CB) between American students and international students. Results indicated no significant difference in Civic Behaviors, *t*(`r round(t_test_cb$parameter, 2)`) = `r round(t_test_cb$statistic, 2)`, *p* = `r round(t_test_cb$p.value, 3)`, with American students (*M* = `r round(mean(civic_data$"Civic Behaviors"[civic_data$"Student Status" %in% c("American", "American.sa")], na.rm = TRUE), 2)`, *SD* = `r round(sd(civic_data$"Civic Behaviors"[civic_data$"Student Status" %in% c("American", "American.sa")], na.rm = TRUE), 2)`) and international students (*M* = `r round(mean(civic_data$"Civic Behaviors"[civic_data$"Student Status" == "International"], na.rm = TRUE), 2)`, *SD* = `r round(sd(civic_data$"Civic Behaviors"[civic_data$"Student Status" == "International"], na.rm = TRUE), 2)`) reporting similar levels. In contrast, a slight but non-significant difference was found for Civic Attitudes, *t*(`r round(t_test_ca$parameter, 2)`) = `r round(t_test_ca$statistic, 2)`, *p* = `r round(t_test_ca$p.value, 3)`, with American students (*M* = `r round(mean(civic_data$"Civic Attitudes"[civic_data$"Student Status" %in% c("American", "American.sa")], na.rm = TRUE), 2)`, *SD* = `r round(sd(civic_data$"Civic Attitudes"[civic_data$"Student Status" %in% c("American", "American.sa")], na.rm = TRUE), 2)`) scoring slightly higher than international students (*M* = `r round(mean(civic_data$"Civic Attitudes"[civic_data$"Student Status" == "International"], na.rm = TRUE), 2)`, *SD* = `r round(sd(civic_data$"Civic Attitudes"[civic_data$"Student Status" == "International"], na.rm = TRUE), 2)`). Overall, these results suggest that while Civic Behaviors and Civic Attitudes collectively do not significantly differ between groups, there is a slight trend in Civic Attitudes favoring American students. These patterns are further illustrated in @fig-civic-behaviors-attitudes, which visualizes the distribution of Civic Behaviors and Civic Attitudes across groups. Additionally, detailed differences in civility towards me and civility towards others was visualized in @fig-civility-me-others. 


## ANOVA
```{r}
#| label: anova-analysis
#| include: false

# Select dependent variables
dependent_vars <- c(
  "Civility Total", 
  "Civility Towards Me",
  "Civility Towards Others",
  "Civic Behaviors", 
  "Civic Attitudes",
  "CB_CA Total", 
  "Wise Reasoning", 
  "Need for Cognition", 
  "Empathy", 
  "Need for Closure", 
  "Epistemic Humility", 
  "Cultural Competence"
)

# Prepare the data for ANOVA
anova_data <- civic_data %>%
  mutate(
    StudentGroup = case_when(
      `Student Status` == "International" ~ "International",
      `Student Status` == "American" ~ "Domestic",
      `Student Status` == "American.sa" ~ "Study Abroad",
      TRUE ~ "Other"
    )
  ) %>%
  filter(StudentGroup != "Other")

# Perform individual ANOVAs for each dependent variable
anova_results <- lapply(dependent_vars, function(var) {
  # Use backticks to handle column names with spaces
  formula <- as.formula(paste("`", var, "` ~ StudentGroup", sep = ""))
  
  # Perform one-way ANOVA
  model <- aov(formula, data = anova_data)
  
  # Get ANOVA summary
  model_summary <- summary(model)[[1]]
  
  # Create a dataframe with results
  result_df <- data.frame(
    Variable = var,
    df_between = model_summary["StudentGroup", "Df"],
    df_within = model_summary["Residuals", "Df"],
    SS_between = model_summary["StudentGroup", "Sum Sq"],
    SS_within = model_summary["Residuals", "Sum Sq"],
    MS_between = model_summary["StudentGroup", "Mean Sq"],
    MS_within = model_summary["Residuals", "Mean Sq"],
    F_value = model_summary["StudentGroup", "F value"],
    p_value = model_summary["StudentGroup", "Pr(>F)"]
  )
  
  return(result_df)
})

# Combine results
anova_summary <- do.call(rbind, anova_results)

# Format p-values
anova_summary$Significance <- ifelse(anova_summary$p_value < 0.001, "***",
                                     ifelse(anova_summary$p_value < 0.01, "**",
                                            ifelse(anova_summary$p_value < 0.05, "*", "ns")))

# Create a nice table of ANOVA results
anova_summary %>%
  kbl(
    caption = "One-way ANOVA Results for Civic Virtues Constructs",
    booktabs = TRUE,
    digits = 3
  ) %>%
  kable_styling(full_width = FALSE)

```

```{r}
#| label: posthoc-analysis
#| include: false

# Perform Tukey's HSD post-hoc test for each significant variable
posthoc_results <- lapply(dependent_vars, function(var) {
  # Use backticks to handle column names with spaces
  formula <- as.formula(paste("`", var, "` ~ StudentGroup", sep = ""))
  
  # Perform Tukey's HSD test
  tukey_result <- TukeyHSD(aov(formula, data = anova_data))
  
  # Convert to data frame
  tukey_df <- as.data.frame(tukey_result$StudentGroup)
  tukey_df$comparison <- rownames(tukey_result$StudentGroup)
  tukey_df$variable <- var
  
  return(tukey_df)
})

# Combine results
posthoc_summary <- do.call(rbind, posthoc_results)

# Create a nice table of all post-hoc results
posthoc_summary %>%
  kbl(
    caption = "Post-hoc Comparisons between Student Groups",
    col.names = c("Difference", "Lower Bound", "Upper Bound", "Adjusted p-value", "Comparison", "Variable"),
    booktabs = TRUE,
    digits = 3
  ) %>%
  kable_styling(full_width = FALSE)
```

A series of one-way analyses of variance (ANOVAs) were conducted to examine potential differences in civic virtue constructs across student groups (International, Domestic, and Study Abroad). The analyses included multiple psychological measures: Civility Total, Civility Towards Me, Civility Towards Others, Civic Behaviors, Civic Attitudes, Wise Reasoning, Need for Cognition, Empathy, Need for Closure, Epistemic Humility, and Cultural Competence.The ANOVA results showed no statistically significant differences between student groups across the examined constructs. Post-hoc Tukey's Honest Significant Difference (HSD) tests further confirmed the lack of significant variations among pairwise group comparisons. Specifically, no significant group differences were observed in any of the measured psychological characteristics. The absence of statistically significant differences suggests that, within this sample, international, domestic, and study abroad students exhibited comparable levels of civic virtues and related psychological attributes. These findings imply that the student experience, regardless of international status, may promote similar developmental trajectories in civic-minded characteristics. The results challenge initial expectations of substantive differences between student groups and highlight the potential homogeneity of civic virtue development in a university setting.

# Discussion
The current study aimed to explore whether international students, due to their unique experiences in a foreign country, exhibit similar or differing levels of civic virtues and related psychological constructs compared to their American peers.
Independent samples t-tests comparing American and international students across key civic virtues revealed no significant differences between the groups on measures of relational civility, civic behaviors and attitudes, wise reasoning, need for cognition, empathy, need for closure, epistemic humility, and cultural competence. These findings suggest that international and American college students in this sample exhibited comparable levels of civic-mindedness and associated psychological characteristics.
The observed similarities in civic virtues among student groups could be attributed to several factors. First, the shared experience of the college admissions process and campus life may create a common context for civic development. Engaging in coursework, extracurricular activities, and community service opportunities provided by the university may equally contribute to students' civic growth, regardless of their international background [@zarrettRoleOrganizedActivities2021]. Second, the self-selection bias inherent in college admissions might result in a student body already predisposed to civic-mindedness, leading to more homogeneous outcomes. 
However, several limitations should be considered when interpreting the results. The sample size, particularly for Amerian students with study abroad experiences, was relatively small, and participants were drawn from a single university. Future research should also employ longitudinal designs to track civic development over time and explore potential mediators and moderators of the relationship between international experiences and civic outcomes. Additionally, the broad categorization of international and American students may obscure important within-group variability. Future studies should consider factors such as English language proficiency and duration of their residence in the United States to better understand the nuances of international students' civic development.

\clearpage

# References

::: {#refs}
:::
